Redshift Evolution of the Underlying Type Ia Supernova Stretch Distribution

### 2020-11-24 Nicolas.tex
All ready for submission, waiting co-authors

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    evolution. In this paper, we study the dependence with redshift of the SN~Ia
    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
    probe its potential redshift drift. The SN stretch has been shown to
    strongly correlate with the SN environment, notably with stellar age
    tracers. We model the underlying stretch distribution as a function of
    redshift, using the evolution of the fraction of young and old SNe~Ia as
    predicted by \cite{rigault2018}, and assuming constant underlying stretch
    distribution for each age population made of Gaussian mixtures. We test our
    prediction against published samples cut to have marginal magnitude
    selection effects, so that any observed change is indeed of astrophysical
    and not observational origin. \nn{In this preliminary study,} there are
    indications that the underlying SN~Ia stretch distribution is evolving as a
    function of redshift, and that the young/old drifting model is a better
    description of the data than any time-constant model, including the
    sample-based asymmetric distributions usually used to correct Malmquist
    bias. The favored underlying stretch model is the bimodal one derived from
    \cite{rigault2018}: a high-stretch mode shared by both young and old
    environments, and a low-stretch mode exclusive to old environments. The
    precise impact of the redshift evolution of the SN Ia population intrinsic
    properties on cosmology remains to be studied. Yet, the astrophysical drift
    of the SN stretch distribution does affect current Malmquist bias
    corrections and thereby distances derived from SN affected by selection
effects. We highlight that such a bias will increase with surveys covering
increasingly larger redshift ranges, which is particularly important for LSST.}

@@

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
and host redshift, the selection effects of magnitude-limited surveys should be
negligible below a given redshift at which even the faintest SNe~Ia can be
observed. In contrast, targeted surveys have highly complex selection functions
and will be discarded from our analysis. Fortunately, modern SN cosmology
samples such as the Pantheon one are now dominated by magnitude-limited surveys.

@@

\nn{We will thus consider a set of cuts that will define a first fiducial
    sample, taking the farthest limits possible; but as this solution might be
    suboptimal to create a complete sample, we also consider another set of cuts
    to define a conservative sample, smaller and therefore less statistically
    constraining, but inherently less prone to selection effects. If the
    redshift drift is still significant in the conservative sample, it would be
    even more significant in a carefully-tailored selection-free sample: the
goal of this preliminary paper is to make a first implementation of this
modeling before refining it on a better sample.}

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement
with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{However,
because this simple approach might easily not be enough to alleviate all
selection effects, we also take into account another, more restrictive,
redshift limit:} Fig.~14 of \citet[][see their Section~5]{perrett2010} suggests a
lower limit of $z_{\lim}=0.55$; both limits will be considered, as discussed
below.

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.31$ in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure is also
suggestive of a more conservative $z_{\lim}=0.27$, \nn{similarly considered in
the following analysis as a first step to evaluate the robustness of our
sample construction.}

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources;
\citet[][see their Section~2]{kessler2009} pointed out that during the first year
of SDSS, SNe~Ia with $r<20.5$~mag were favored for spectroscopic follow-up,
corresponding to a redshift cut at $0.15$. For the rest of the SDSS survey,
additional spectroscopic resources were available, and \cite{kessler2009} and
\cite{dilday2008} show a reasonable completeness up to $z_{\lim}=0.2$. Following
these analyses, we will use $z_{\lim}=0.20$ as the baseline SDSS redshift limit,
\nn{and $z_{\lim}=0.15$ to compare the results of our modeling on a more
restrictive sample.}

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits roughly correspond to the peak of these
histograms. \nn{We test in Section~\ref{ssec:verify} whether we can consider
this sample to be free of (or at least marginally affected by) selection
effects. The consistency of the results for the two samples, discussed} in
Section~\ref{sec:results}, further supports that this sample selection indeed
provides a subset of SNe~Ia with insignificant selection effects when compared
to state-of-the-art Malmquist correction techniques.

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). The spectroscopic follow-up was done over a redshift
range of $0.02 < z < 0.09$~\citep[as in][]{rigault2018}, while the search was
much deeper. These 114 SNfactory SNe~Ia \nn{(80\% of the 141 SNe presented in
the previously-cited paper)} are thus in the volume-limited part of the
survey (Aldering et al., in prep.), and are therefore assumed to be a random
sampling of the underlying SN population. The SNfactory sample is
particularly useful for studying SN property drift, as it enables us to have
a large complete SN~Ia sample at $z<0.1$.

Finally, we include the HST sample from Pantheon, that similarly have a search
deeper than the follow-up and that we therefore kept entirely
\citep{strolger04}.

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \nn{Proportionally
few low-stretch SNe (typically $x_1 < -1$) are observed at $z>0.2$}. The
evolution of the mean stretch is also shown in Fig.~\ref{fig:modelall}, where
the data are split in redshift bins of regular sample size. We see that SNe~Ia
at higher redshift have on average larger stretch ($0.34 \pm 0.10$ at
$z\sim0.65$) than those at lower redshift ($-0.17\pm 0.10$ at $z\sim0.05$),
suggesting that the underlying stretch distribution is evolving with redshift.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_stac_75-lb-cividis.pdf}
    \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} lightcurve stretch as a
        function of redshift for each survey considered in this analysis (see
        legend). Solid (resp.\ open) markers correspond to the conservative
        (resp.\ fiducial) redshift cuts. \textit{Top:} stacked redshift
    histograms in dark (resp.\ light) colors for the conservative (resp.  \
fiducial) redshift cuts. \yc{YC: you should add a running mean $\pm$ stddev to
show the evolution at $z \sim 0.2-0.3$.}}
    \label{fig:sample}
\end{figure}

@@

In this analysis, we have built a volume-limited sample from a set of magnitude
limited ones, using a simple redshift limit. \nn{This drastic approach is robust
at the price of being statistically sub-optimal, but is enough to
demonstrate our key finding: a redshift evolution of stretch compatible with
the \cite{rigault2018} model.} However, while we will find consistent results
when considering a sample derived from conservative cuts
(Section~\ref{sec:results}), there remains the possibility that our sample is
not fully volume-limited and that unaccounted for selection effects remain. To
fully account for this possibility, a follow up analysis will make use of the
SNANA package \citep{SNANA} to account for more subtle selection functions;
\nn{yet, it appears to still be necessary} to test if the sample we have created
could show sign of leftover sample selection.

To do so, we compare the stretch and color distributions of the SNe~Ia in our
sample between datasets of overlapping redshifts. If indeed we have built a
volume-limited sample, the distributions of any SNe~Ia parameter at a given
redshift should not depend of their survey of origin. When considering a
redshift range, this has to be narrow enough such that any drift would be
negligible. \yc{YC: \textit{why do you say that here, since you didn't fine-tune
the overlaping $z$-range?}} \nn{\textit{Because we still took that into
account?}} The two samples that overlap the most in redshift are PS1 and SDSS in
the redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This
corresponds to the 146 SNe~Ia from SDSS at the highest redshift end and thus the
most affected by potential selection effect leftover (see the corresponding
discussion in section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52
SNe~Ia, which are in the lowest redshift bins and thus unlikely to have any
selection issue. Fig.~\ref{fig:distrib} presents the stretch and color
distribution of both \nn{these surveys}, together with $p$-values of
Kolmogorov-Smirnov tests that \yc{\sout{\nn{probe whether}}}\nn{\textit{I don't
see anything else to put here without repeating ``test''}} the $x_1$ and $c$ of
the SDSS and PS1 samples at $0.10 < z < 0.20$ arise from the same underlying
distribution. These $p$-values are high ($>10\%$), and if one cannot conclude
from this test that these SNe~Ia are from the same underlying distribution, one
has no ground to claim they are not. \yc{YC: I suggest ``These $p$-values are
too high ($>10\%$) to exclude that both samples originate from the same
underlying population.''} This is in agreement with the visual impression from
Fig.~\ref{fig:distrib}. We perform a similar analysis for PS1 and SNLS in the
redshift range $0.20 < z < 0.31$ where the same conclusion can be drawn, i.e.,
there is no sign of discrepancy in the stretch and color distributions between
the low- and high-end of the SNLS and PS1 datasets, respectively. Nonetheless,
the small size of the SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1)
limits the interest of the test as only strong deviation would be noticeable.
Extending the redshift range to $0.20 < z < 0.40$ (though we have no PS1 data
above 0.3) enables to get 31 SNe~Ia from SNLS in total, yet the stretch and
color KS-test $p$-values remains inconclusive, at 34\% and 4\% respectively.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
    \caption{$x_1$ (left) and $c$ (right) distribution histograms of different
        surveys overlapping in redshift. \textit{Facing up}: SDSS and PS1 within
        the redshift range $0.10 < z < 0.20$; \textit{facing down}: PS1 and SNLS
        within the redshift range $0.20 < z < 0.31$. Error bars show the Poisson
        noise \yc{YC: not visible enough}. Our stretch ``Base'' model is
        illustrated in orange at the mean redshift of the redshift ranges,
        $0.15$ and $0.25$, respectively.  Kolmogorov-Smirnov test $p$-values are
    indicated on the top (resp. bottom) of each panel showing no sign that the
SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$ distributions are not drawn from
the same underlying distributions.}
    \label{fig:distrib}
\end{figure}

We furthermore highlight that the SNe~Ia color is more prone to selection
effects than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g.,
Fig.~3 of \cite{kessler2017}. Hence, since the color distribution shows no
significant hint of selection effect, it further supports our claim that our
simple redshift-based selection criteria is sufficient to \nn{build an unbiased
SNe~Ia sample at all redshifts required to test the redshift evolution of the
stretch distribution}.

@@

We report in Table~\ref{tab:bbc} \nn{our determination of} the samples' $\mu_0$
and $\sigma_{\pm}$ \nn{when implementing an asymmetric Gaussian model}, adjusted
on the nominally selection-free samples using our fiducial cuts (see
Section~\ref{sec:sample}). We find our results in close agreement with
\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
derived these model parameters using the full BBC formalism, using numerous
simulations to model the selection effects (see details e.g., Section~3 of
\citealt{kessler2017}). The agreement between our fit of the asymmetric
Gaussians on the supposedly selection-free part of the samples and the results
derived using the BBC formalism supports our approach to get a sample with
negligible selection effects. If we were to use \cite{scolnic2016} and
\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
Base drifting model and the BBC modeling would go even deeper from $-32$ to
$-47$. We further discuss the consequence of this result for cosmology in
Section~\ref{sec:discussion}.

@@

We also performed tests allowing the high-stretch mode of the old population to
differ from the young population mode, hence adding two degrees of freedom. The
corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
this strengthens the fact that the young and old populations indeed appear to
share the same underlying high-stretch mode. Furthermore, one might wonder
whether a low-stretch mode might also exist in the young-population, see
Fig.~\ref{fig:stretchlssfr}. We tested that by allowing this population to also
be bi-modal, finding the amplitude of this low-stretch mode to be compatible
with~0 in this young population ($<2\%$). More generally, this raises the
question of inaccurate tracing of age by a given environmental tracer (here the
LsSFR). A dedicated analysis will be presented in Briday et al.\ in prep.

@@

We have presented a study of the drift of the underlying SNe~Ia stretch
distribution as a function of redshift. We built a magnitude-limited SN~Ia
sample from the Pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a}, to
which we added HST and SNfactory data from \cite{rigault2018} for the high- and
low-redshift bins. We only considered the SNe that have been discovered in the
redshift range of each survey where selection effects are negligible, so that
the observed SNe~Ia stretches are random sampling of the true underlying
distribution. This resulted in a 569 SN~Ia fiducial sample (422 SNe when more
conservative cuts were considered).

@@

Our conclusions are the following:
\begin{enumerate}
    \item The underlying SN~Ia stretch distribution is significantly redshift
        dependent, as previously suggested by e.g.~\cite{howell2007} \nn{in a
        way that selection effects alone cannot explain}. This result is largely
        independent of details on each age-population modeling.

@@

In this paper, we considered a simple two-population Gaussian mixture modeling,
but additional data free from significant Malmquist bias would enable us to
refine it as necessary. We note that samples at the low- and high-redshift ends
of the Hubble diagram would be particularly helpful for this drifting analysis;
fortunately this will soon be provided by the Zwicky Transient Facility
\citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange SNe~Ia
programs (high-$z$), respectively. \nn{To improve the disentangling of
observational} selection effects and \nn{intrinsic} population evolution, an
analysis using the SNANA framework is underway. It will allow for subtle
selection effects to be \nn{accounted for}, and by removing the need for a
\nn{naive} redshift cut, increase the redshift leverage and statistical
significance of our results.

@@

\yc{YC: check again all arXiv references.}

@@

\bibitem[Freedman et al.(2020)]{freedman2020} Freedman, W.~L., Madore, B.~F.,
Hoyt, T., et al.\ 2020, arXiv e-prints, arXiv:2002.01550

@@

\bibitem[Knox \& Millea(2019)]{knox2019} Knox, L., \& Millea, M.\ 2019, arXiv
e-prints, arXiv:1908.03663

@@

\bibitem[Planck Collaboration et al.(2018)]{planck2018} Planck Collaboration,
Aghanim, N., Akrami, Y., et al.\ 2018, arXiv e-prints, arXiv:1807.06209

@@

\bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess,
A.~G.\ 2019, arXiv e-prints, arXiv:1908.05625

@@

\bibitem[Wiseman et al.(2020)]{wiseman2020} Wiseman, P., Smith, M., Childress,
M., et al.\ 2020, arXiv e-prints, arXiv:2001.02640

\bibitem[Wong et al.(2019)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F.,
et al.\ 2019, arXiv e-prints, arXiv:1907.04869

>>>>

\abstract{The true nature of type Ia supernovae (SNe~Ia) remains largely
    unknown, and as survey statistics increase, the question of astrophysical
    systematic uncertainties rises, notably that of the SN~Ia population
    evolution. In this paper, we study the dependence with redshift of the SN~Ia
    \texttt{SALT2.4} lightcurve stretch, a purely intrinsic SN property, to
    probe its potential redshift drift. The SN stretch has been shown to
    strongly correlate with the SN environment, notably with stellar age
    tracers. We model the underlying stretch distribution as a function of
    redshift, using the evolution of the fraction of young and old SNe~Ia as
    predicted by \cite{rigault2018}, and assuming constant underlying stretch
    distribution for each age population made of Gaussian mixtures. We test our
    prediction against published samples \textbf{cut} to have marginal magnitude
    selection effects, so that any observed change is indeed of astrophysical
    and not observational origin. \textbf{In this first study, there are
    indications} that the underlying SN~Ia stretch distribution is evolving as a
    function of redshift, and that the young/old drifting model is a better
    description of the data than any time-constant model, including the
    sample-based asymmetric distributions usually used to correct Malmquist
    bias. The favored underlying stretch model is the bimodal one derived from
    \cite{rigault2018}: a high-stretch mode shared by both young and old
    environments, and a low-stretch mode exclusive to old environments. The
    precise impact of the redshift evolution of the SN Ia population intrinsic
    properties on cosmology remains to be studied. Yet, the astrophysical drift
    of the SN stretch distribution does affect current Malmquist bias
    corrections and thereby distances derived from SN affected by selection
effects. We highlight that such a bias will increase with surveys covering
increasingly larger redshift ranges, which is particularly important for LSST.}

@@

Assuming sufficient (and unbiased) spectroscopic follow-up for acquiring typing
and host redshift, the selection effects of magnitude-limited surveys should be
negligible below a given redshift at which even the faintest SNe~Ia can be
observed. \textbf{In practice such spectroscopic following has not always been
the norm as discussed below.} In contrast, targeted surveys have highly
complex selection functions and will be discarded from our analysis.
Fortunately, modern SN cosmology samples such as the Pantheon one are now
dominated by magnitude-limited surveys.

@@

\textbf{We will thus consider a set of cuts that will define a first fiducial
sample, taking the farthest limits possible. But as this solution might be
over simplistic to create a complete sample, e.g., for it ignores follow-up
efficiency, we also consider another set of cuts to define a conservative
sample. This latter is smaller and therefore less statistically
constraining, but also inherently less prone to selection effects. If the
redshift drift is still significant in the conservative sample, it would be
even more significant in a carefully-tailored selection-free sample. The
goal of this study is to build the first implementation of a SN Ia
properties drift modeling before refining it on a more complex, yet more
accurate, follow-up analysis accounting at the same time for precise selection
modeling using of the SNANA package \citep{SNANA}.}

SNLS typically acquired SNe~Ia in the redshift range $0.4<z<0.8$; at these
redshifts, the rest-frame Bessel $B$ band roughly corresponds to the SNLS $i$
filter, that has a $5\sigma$ depth of
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{\lim}=0.60$, in agreement with
\cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. Fig.~14 of
\citet[][see their Section~5]{perrett2010} suggests however a lower limit of
$z_{\lim}=0.55$. \textbf{We will therefore use $z=0.60$ and $z=0.55$ as redshift
limits for the fiducial and conservative cuts, respectively.}

Similarly, PS1 observed SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1~mag \citep{rest2014}, which yields to $z_{\lim}=0.31$ in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. \textbf{If we were to be
conservative, this figure would also suggest of a more stringent
$z_{\lim}=0.27$ cut; hence, we will use $0.31$ and $0.27$ for our fiducial and
conservative cuts, respectively.}

In a similar redshift range, SDSS has a limiting magnitude of 22.5
\citep{dilday2008, sako2008}, which would lead to a $z_{\lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources
\textbf{as discussed before}. \citet[][Section~2]{kessler2009} pointed out that
during the first year of SDSS, SNe~Ia with $r<20.5$~mag were favored for
spectroscopic follow-up, corresponding to a redshift cut at $0.15$. For the rest
of the SDSS survey, additional spectroscopic resources were available, and
\cite{kessler2009} and \cite{dilday2008} show a reasonable completeness up to
$z_{\lim}=0.2$. Following these analyses, \textbf{we will use $z_{\lim}=0.20$
and $z_{\lim}=0.15$ for out fiducial and conservative cuts, respectively.}

The sample selection is summarized in Table~\ref{tab:sample}, and the redshift
distribution of these three surveys is shown in Fig.~\ref{fig:cuts}. As
expected, the selected redshift limits \textbf{are roughly located slightly
before the peak of these histograms}. \textbf{We test in
Section~\ref{ssec:verify} the validity of our intent to build volume-limited
surveys from (mostly) magnitude-limited ones with the aforementioned redshift
limits.}

@@

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory,][]{aldering2002} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their sections~3
and~4.2.2; SNe~Ia time series are published in \citealt{saunders2020}, see also
\citealt{aldering2020}). \textbf{For this dataset,} the spectroscopic follow-up
was done over a redshift range of $0.02 < z < 0.09$, while the search was much
deeper. These 114 SNfactory SNe~Ia are thus in the volume-limited part of the
survey (Aldering et al., in prep.), and are therefore assumed to be a random
sampling of the underlying SN population. The SNfactory sample is particularly
useful for studying SN property drift, as it enables us to have a large complete
SN~Ia sample at $z<0.1$.

Finally, we include the HST sample from Pantheon, that similarly have a search
deeper than the follow-up and that we therefore kept entirely
\citep{strolger04}. \textbf{At these redshift, the supernovae typing is
challenging but the target classification were robust enough to include them
within the cosmological analysis \citep{scolnic2018a}; we do not add further
quality cuts.}

We present the stretch distribution and redshift histogram of these five surveys
up to their respective $z_{\lim}$ in Fig.~\ref{fig:sample}. \textbf{We observe
that the apparent fraction of low-stretch SNe (typically $x_1 < -1$) seems to
decrease as a function of redshift.} The evolution of the mean stretch is also
shown in Fig.~\ref{fig:modelall}, where the data are split in redshift bins of
regular sample size. We see that SNe~Ia at higher redshift have on average
larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower redshift
($-0.17\pm 0.10$ at $z\sim0.05$), suggesting that the underlying stretch
distribution is \textbf{evolving with redshift}.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_stac_75-lb-cividis.pdf}
    \caption{\textit{Bottom:} \textsc{\texttt{SALT2.4}} lightcurve stretch as a
        function of redshift for each survey considered in this analysis (see
        legend). Solid (resp.\ open) markers correspond to the conservative
        (resp.\ fiducial) redshift cuts. \textit{Top:} stacked redshift
    histograms in dark (resp.\ light) colors for the conservative (resp.  \
fiducial) redshift cuts.}
    \label{fig:sample}
\end{figure}

@@

\textbf{In section~\ref{ssec:cuts}, we have built a volume-limited sample from a
set of magnitude-limited ones, using a simple redshift limit. This
simplistic approach is statistically sub-optimal, but is enough to
demonstrate our key finding: a redshift evolution of stretch compatible with
the \cite{rigault2018} model. However, there remains the possibility that
complex selection function related to follow-up efficiencies below our fiducial
(or even conservative cuts) might significantly affect our sample, making it not
fully volume-limited, which could, in turn, bias our conclusion on the
astrophysical drift of the SNe~Ia population.}

\textbf{To test for the existence of potentially significant leftover selection biases in our sample, we compare the stretch and color distributions of the SNe~Ia in our
sample between datasets of overlapping redshifts. If indeed we have built a
volume-limited sample, the distributions of any SNe~Ia parameter at a given
redshift should not depend of their survey of origin. When considering a
redshift range, this has to be narrow enough such that any drift would be
negligible.}

%\yc{YC: \textit{why do you say that here, since you didn't fine-tune
%the overlaping $z$-range?}} \nn{\textit{NN: Because we still took that into
%account?}} \mr{\textit{MR:Because if there is a strong drift in the redshift
%range, and if the redshift distribution of the sample differ, their respective
%property distribution should differ.}}

\textbf{The two samples that overlap the most in redshift are PS1 and SDSS in the
redshift range $0.10 < z < 0.20$ (see Fig~\ref{fig:sample}). This corresponds to
the 146 SNe~Ia from SDSS at the highest redshift end and thus the most affected
by potential selection effect leftover (see the corresponding discussion in
section~\ref{ssec:cuts}). In that same redshift range, PS1 has 52 SNe~Ia, which
are in the lowest redshift bins and thus unlikely to have any selection issue.
Fig.~\ref{fig:distrib} presents the stretch and color distribution of both these
surveys, together with $p$-values of Kolmogorov-Smirnov tests (KS-test). These
$p$-values are too high ($>10\%$) to exclude that both samples originate from
the same underlying population, in agreement with the visual impression from
Fig.~\ref{fig:distrib}.}

%The $p$-values are high ($>10\%$), and if one cannot conclude from a KS-test
%that indeed these SNe~Ia are from the same underlying distribution, there is no
%ground to claim they are not.

\textbf{We perform a similar analysis for PS1 and SNLS in the redshift range $0.20 < z <
0.31$ where the same conclusion can be drawn, i.e., there is no sign of
discrepancy in the stretch and color distributions between the low- and high-end
of the SNLS and PS1 datasets, respectively. Nonetheless, the small size of the
SNLS dataset at $z < 0.31$ (12 SNe~Ia vs. 90 for PS1) limits the interest of the
test as only strong deviation would be noticeable.  Extending the redshift range
to $0.20 < z < 0.40$ (though we have no PS1 data above 0.3) enables to get 31
SNe~Ia from SNLS in total, yet the stretch and color KS-test $p$-values remains
inconclusive, at 34\% and 4\% respectively.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/both-cut_SDSS_SNLS_PS1.pdf}
    \caption{\textbf{$x_1$ (left) and $c$ (right) distribution histograms of
            different surveys overlapping in redshift. \textit{Facing up}: SDSS
            and PS1 within the redshift range $0.10 < z < 0.20$; \textit{facing
            down}: PS1 and SNLS within the redshift range $0.20 < z < 0.31$.
            Error bars show the Poisson noise. Our stretch ``Base'' model is
            illustrated in orange at the mean redshift of the redshift ranges,
            $0.15$ and $0.25$, respectively.  Kolmogorov-Smirnov test $p$-values
            are indicated on the top (resp.  bottom) of each panel showing no
            sign that the SDSS and PS1 (resp. PS1 and SNLS) $x_1$ and $c$
            distributions are not drawn from the same underlying
    distributions.}}
    \label{fig:distrib}
\end{figure}

\textbf{We highlight that the SNe~Ia color is more prone to selection effects
than stretch as illustrated in Fig.~\ref{fig:maglim}; see also e.g., Fig.~3
of \cite{kessler2017}. Hence, since the color distribution shows no
significant hint of leftover selection effect, it further supports our claim
that our simple redshift-based selection criteria is sufficient to build the
SNe~Ia sample required to test the redshift evolution of the stretch
distribution.}

@@

We report in Table~\ref{tab:bbc} \textbf{our determination of} the samples'
$\mu_0$ and $\sigma_{\pm}$ \textbf{when implementing an asymmetric Gaussian
model}, adjusted on the nominally selection-free samples using our fiducial cuts
(see Section~\ref{sec:sample}). We find our results in close agreement with
\cite{scolnic2016} for SNLS and SDSS and with \cite{scolnic2018a} for PS1, who
derived these model parameters using the full BBC formalism, using numerous
simulations to model the selection effects (see details e.g., Section~3 of
\citealt{kessler2017}). The agreement between our fit of the asymmetric
Gaussians on the supposedly selection-free part of the samples and the results
derived using the BBC formalism supports our approach to get a sample with
negligible selection effects. If we were to use \cite{scolnic2016} and
\cite{scolnic2018a} best fit values of the $\mu_0, \sigma_{\pm}$ asymmetric
parameters for SNLS, SDSS and PS1, respectively, the $\Delta$AIC between our
Base drifting model and the BBC modeling would go even deeper from $-32$ to
$-47$. We further discuss the consequence of this result for cosmology in
Section~\ref{sec:discussion}.

@@

We also performed tests allowing the high-stretch mode of the old population to
differ from the young population mode, hence adding two degrees of freedom. The
corresponding fit is not significantly better, with a $\Delta$AIC of $-0.4$.
This strengthens the fact that the young and old populations indeed appear to
share the same underlying high-stretch mode. Furthermore, one might wonder
whether a low-stretch mode might also exist in the young-population, see
Fig.~\ref{fig:stretchlssfr}. We tested that by allowing this population to also
be bi-modal, finding the amplitude of this low-stretch mode to be compatible
with~0 in this young population ($<2\%$). More generally, this raises the
question of inaccurate tracing of age by a given environmental tracer (here the
LsSFR). A dedicated analysis will be presented in Briday et al.\ in prep.

@@

We have presented a \textbf{first} study of the drift of the underlying SNe~Ia
stretch distribution as a function of redshift. We built a magnitude-limited
SN~Ia sample from the Pantheon dataset \citep[][SDSS, PS1 and
SNLS]{scolnic2018a}, to which we added HST and SNfactory data from
\cite{rigault2018} for the high- and low-redshift bins. We only considered the
SNe that have been discovered in the redshift range of each survey where
selection effects are negligible, so that the observed SNe~Ia stretches are
random sampling of the true underlying distribution. This resulted in a 569
SN~Ia fiducial sample (422 SNe when more conservative cuts were considered).

@@

Our conclusions are the following:
\begin{enumerate}
    \item The underlying SN~Ia stretch distribution is significantly redshift
        dependent, as previously suggested by e.g.~\cite{howell2007} \textbf{in
        a way that selection effects alone cannot explain}. This result is
        largely independent of details on each age-population modeling.

@@

In this paper, we considered a simple \textbf{two-population} Gaussian mixture
modeling, but additional data free from significant Malmquist bias would enable
us to refine it as necessary. We note that samples at the low- and high-redshift
ends of the Hubble diagram would be particularly helpful for this drifting
analysis; fortunately this will soon be provided by the Zwicky Transient
Facility \citep[low-$z$,][]{bellm2019, graham2019}, and Subaru and SeeChange
SNe~Ia programs (high-$z$), respectively.

\textbf{The next step of the analysis will be to incorporate our modeling within
the SNANA framework \citep{SNANA} both to more accurately account for selection
functions and to test the impact of our modeling on the derivation of
cosmological parameters; this study is underway.}

@@

@@

\bibitem[Freedman et al.(2020)]{freedman2020} Freedman, W.~L., Madore, B.~F.,
Hoyt, T., et al.\ 2020, \apj, 891, 57. doi:10.3847/1538-4357/ab7339

@@

\bibitem[Knox \& Millea(2020)]{knox2019} Knox, L. \& Millea, M.\ 2020, \prd,
101, 043533. doi:10.1103/PhysRevD.101.043533

@@

\bibitem[Planck Collaboration et al.(2020)]{planck2018} Planck
Collaboration, Aghanim, N., Akrami, Y., et al.\ 2020, \aap, 641, A6.
doi:10.1051/0004-6361/201833910

@@

\bibitem[Reid et al.(2019)]{reid2019} Reid, M.~J., Pesce, D.~W., \& Riess,
A.~G.\ 2019, \apjl, 886, L27. doi:10.3847/2041-8213/ab552d

@@

\bibitem[Wiseman et al.(2020)]{wiseman2020} Wiseman, P., Smith, M., Childress,
M., et al.\ 2020, \mnras, 495, 4040. doi:10.1093/mnras/staa1302

\bibitem[Wong et al.(2020)]{wong2019} Wong, K.~C., Suyu, S.~H., Chen, G.~C.-F.,
et al.\ 2020, \mnras, 498, 1420. doi:10.1093/mnras/stz3094
